|

### 2022-06-22 docs/rmd/paper.Rmd
autogit

---
title: |
  Analyzing the influence of Selection on Genetic Programming's Generalization ability in Symbolic Regression:
    A comparison of $\epsilon$-Lexicase Selection and Tournament Selection
date: "Date of Submission: `r Sys.Date()`"
output:
  word_document:
    reference_docx: "templates/template_ls.docx"
    fig_caption: yes
    number_sections: true
always_allow_html: true 
fontsize: 12pt
toc: yes
toc_depth: 4
bibliography: bib/bib/ref.bib
csl: bib/csl/harvard1.csl
include-before: '`\newpage{}`{=latex}'
---

```{r setup, include=FALSE}

library(kableExtra)
library(magrittr)

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.pos = "!H")

# Abstract

...

@@

The basic algorithm for lexicase parent selection as described by @lexicase_first_desciption is given below:

@@

## Descriptive Statistics

Tables 6 and 7 summarize the results for all fitness scores collected over 50 total runs of the experiment^[All floating point numbers in text have been rounded to 3 decimal places].

@@

<!-- Statistical Analysis

For both experiments the best of the run models that are evolved using $\epsilon$-Lexicase Selection on average perform better as measured by their mean error score. This behaviour is consistent during both training and testing phases of the experiment

$\epsilon$-Lexicase based GP achieves an average minimum fitness of 3.166/1.576 (MSE/MAE) during model training and a score of 2.949/1.678 during the testing phase on previously unseen data. Suprisingly the models that are trained using MSE as a metric for fitness evaluation, on average, perform even better on unseen cases as they do on the initial training data.

In comparison, the models that are trained using tournament selection perform with an average minimum fitness of 10.9887/2.32173 (MSE/MAE) during model training and 11.146/2.196 during the testing phase. Again we observe an improved performance in testing error, this time models that are evolved using MAE for fitness perform better on testing data than on their initial training data.

The initial results seem support the two alternative hypothesis of this research project:

- $HA_{testing}$: T $\epsilon$-Lexicase Selection produces models that differ in their generalization behaviour if compared to models that are evolved using torunament selection

- $H0_{training}$: $\epsilon$-Lexicase Selection produces models that differ in their performance during model training if compared to models that are evolved using torunament selection

--->

## Inferential Statistics

Figure 1 visualizes the distribution of fitness scores achieved by both GP systems after finishing 100 full generations for all 50 runs of the experiment.

@@

The four samples from figure 1 have been tested for normal distribution by computing the D’Agostino and Pearson test [@DAgostino1971AnOT] [@DAgostino1973TestsFD] to determine if they satisfy the requirements of the students t-test. The results are summarized in table 8. Since not all samples test positive for a normal distribution a Mann-Whitney U (MWU) ranksum test will be performed to test for statistical significance.

@@

The results of the MWU test are summarized in table 9.

@@

@@

>>>>

---
title: |
  <center>Analyzing the influence of Selection on Genetic Programming's Generalization ability in Symbolic Regression: </center>
  <center>
    A comparison of $\epsilon$-Lexicase Selection and Tournament Selection
  </center>
  <center>--------------------------------------------------------</center>
  <center>Student: Roman Höhn</center>
  <center>Date of Birth: 1991-04-14</center>
  <center>Place of Birth: Wiesbaden, Hesse</center>
  <center>Student ID: 2712497</center>
  <center>Supervisor: David Wittenberg</center>
  <center>--------------------------------------------------------</center>
  <center>Research Project for Seminar Information Systems (03.996.3299)</center>
  <center>Chair of Business Administration and Computer Science</center>
  <center>Johannes Gutenberg University Mainz</center>
  <center>Summerterm 2022</center>
  |
date: "Date of Submission: `r Sys.Date()`"
output:
  pdf_document:
    keep_tex: true
    extra_dependencies: ["float"]
    fig_caption: yes
    number_sections: true
    includes:
      in_header: settings.tex
fontsize: 12pt
# toc: yes
# toc_depth: 4
bibliography: bib/bib/ref.bib
csl: bib/csl/harvard1.csl
# include-before: '`\newpage{} `{=latex}'
include-before: '`\thispagestyle{empty} `{=latex}'
---

```{r setup, include=FALSE}
library(kableExtra)
library(magrittr)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.pos = "!H")
```

\tableofcontents
\thispagestyle{empty}
\newpage

@@

The first description^[@lexicase_first_desciption, p.4] of the lexicase parent selection algorithm by @lexicase_first_desciption is quoted below for reference:

@@

Tables 6 and 7 summarize the results for all fitness scores collected over 50 total runs of the experiment^[All floating point numbers in tables have been rounded to 3 decimal places if not stated otherwise].

@@

Some interesting observation from tables 6 and 7 can be made by comparing the mean of the variables $min\_training\_error$ and $testing\_error$. Variable $min\_training\_error$ represents the fitness, measured by the MSE, of the best performing solution that has been found during a single run of the experiment while the variable $testing\_error$ represents the fitness of the same solution if evaluated for the unknown fitness cases inside the testing dataset. Tournament selection based GP achieves a mean $min\_training\_error$ of $16.483$ and a mean $testing\_error$ of $16.905$ while the GP system using $\epsilon$-Lexicase selection can only achieve a mean $min\_training\_error$ of $22.013$ and a mean $testing\_error$ of $22.456$. These results indicate an overall performance benefit for torunament selection based GP in the experiment. It should also be mentioned that the best performing model over all 50 runs, measured by the minimum of variables $min\_training\_error$ and $testing\_error$, has been achieved by tournament selection with an MSE of 7.066 for the training data and	7.847 for the testing data.

If we compare the differences of $min\_training\_error$ and $testing\_error$ on the basis of each algorithm, we can observe that both tournament and $\epsilon$-lexicase achieve a slightly lower error on the training data than on the testing data. If we compute the total difference between the means of $min\_training\_error$ and $testing\_error$ for both algorithms tournament selection amounts to 0.422 and $\epsilon$-Lexicase selection to 0.443. Regarding our primary research question, the differences in generalization, the initial observations do not seem to indicate a significant difference in the the relative gap between the solutions performance on testing data and training data.

Figure 1 visualizes the distribution of $min\_training\_error$ and $testing\_error$ for both selection operators over the total 50 runs using a boxplot and summerizes the initial observations from tables 6 and 7 described above.

@@

The four samples from figure 1 have been tested for normal distribution by computing the D’Agostino and Pearson test [@DAgostino1971AnOT] [@DAgostino1973TestsFD] to determine if they satisfy the requirements of the students t-test. The results are summarized in table 8. Since not all samples test positive for a normal distribution a Mann-Whitney U (MWU) ranksum test will be performed to test for statistical significance between the four variables $tournament\_min\_training\_error$, $tournament\_testing\_error$, $e-lexicase\_min\_training\_error$ and $e-lexicase\_testing\_error$.

@@

Table 9 displays the resulting P-values from computing the MWU test in a matrix format. The results demonstrate a highly significant difference int

@@

df <- read.csv("./tables/csv/mwu_matrix_error.csv")

@@

The mean overall performance of tournament selection appears to be superior to that of $\epsilon$-Lexicase selection which could be considered contradictory evidence to the 
... interpretation...
